GPU Virtualization in the Cloud (a mejorar el t\'itulo!!!)

### 2015-10-05 paper.tex
Titulo modificado
experimental OS

\noindent 
Nowadays, to provide a virtual machine with GPUs means
that the guest host must be equipped with at least 1 GPU, and
once it has been assigned to an instance, no other VM can use
it. Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
the conservative, to have a lot of hosts with GPUs to be
ensured that you are likely to satisfy all the requests; or a
more daring approach, where you will count with some GPU-
enabled nodes hoping not to arrive a large burst of requests
so the cluster would run out of GPU resources. Our solutions
is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand.
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project. On
the one hand the work done in [1] allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node. With the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines. The solution presented
by the project gVirtus [2] does virtualize GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in
[3] by the name of gCloud. This one continues without being
integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. On the
other hand, the work exposed in [4] is more mature, however,
it is only focussed on compute intensive HPC applications.
The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

@@

\section{Motivation}
\label{sec:motivation}

@@

\section{GPGPU Service for OpenStack}
\subsection{Managing remote GPUs with OpenStack}
The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part into the OpenStack Architecure.
This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
As it is shown in Figure~\ref{fig:internal}, we have altered the original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
Thanks to this modular approach, the Compute Module has not been modified, which led into an easily portability to other Cloud Computing environments.

@@

\subsection{Development}
We have developed an independent module that can be seen as a new shared service in the OpenStack Architecture (see Figure~\ref{fig1}).

Thus, we have provided OpenStack with a new parser, which splits the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
Thanks to this modular approach, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.

The new GPGPU Service has brought a new feature when it comes to GPUs, from now on, we can create ``GPU-pools''. A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

Furthermore, we have extended the user interface (Horizon Project), in order to deal with the new features.
First of all, the Instance Launch Panel has been provided with a new field, where the user can assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' has been chosen, will appear the proper fields for the pool configuration.
Finally, with the purpose of being aware of the devices status, a new panel has been added with all the information regarding the GPUs.

@@

\subsection{User Interface}
Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
First of all, the Instance Launch Panel has been provided with a new field, where the user can assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' has been chosen, will appear the proper fields for the pool configuration (see Subfigure~\ref{fig:ui-launch}).
Furthermore, a new panel with the existent GPUs has been added to the Interface (see Subfigure~\ref{fig:ui-rgpus}), where all the information of the GPUs is displayed.

\begin{figure}[!t]
  \centering
  \includegraphics[width=.9\linewidth]{images/UI-launch.jpg}
  \caption{Launching Instances and assigning GPUs}
  \label{fig:ui-launch}
\end{figure}

\begin{figure}[!t]
  \centering
  \includegraphics[width=.85\linewidth]{images/UI-rgpus.jpg}
  \caption{GPU Information Panel}
  \label{fig:ui-rgpus}
\end{figure}

\subsection{Experimental Results}

\section{Discussion}
\label{sec:discussion}
Volver a hacer un resumen

>>>>

\noindent

@@

\section{Motivation}
\label{sec:motivation}
Nowadays, to provide a virtual machine with GPUs means
that the guest host must be equipped with at least a GPU, and
once it has been assigned to an instance, no other VM can use
it. Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
the conservative, to have a lot of hosts with GPUs to be
ensured that you are likely to satisfy all the requests; or a
more daring approach, where you will count with some GPU-
enabled nodes hoping not to arrive a large burst of requests
so the cluster would run out of GPU resources.

Our solutions
is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand.
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project. On
the one hand the work done in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node. With the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines. The solution presented
by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. On the
other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute intensive HPC applications.
The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

@@

\section{GPGPU Service for OpenStack}
\subsection{Managing remote GPUs with OpenStack}
Our contribution reaches a solution that drive us to a scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part into the OpenStack Architecure.
This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

@@

Thanks to the modular approach of the GPGPU Service, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.

@@

\subsection{User Interface}
Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
First of all, the Instance Launch Panel should be provided with a new field, where the user could assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' had chosen, would appear the proper fields for the pool configuration (see Figure~\ref{fig:ui-launch}).
Furthermore, a new panel with the existent GPUs would display all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).

\begin{figure}[!t]
  \centering
  \includegraphics[width=\linewidth]{images/UI-launch.jpg}
  \caption{Launching Instances and assigning GPUs}
  \label{fig:ui-launch}
\end{figure}

\begin{figure}[!t]
  \centering
  \includegraphics[width=\linewidth]{images/UI-rgpus.jpg}
  \caption{GPU Information Panel}
  \label{fig:ui-rgpus}
\end{figure}

\subsection{Experimental Results}
The test were executed on a 5-node cluster where a machine
equipped with a Intel Xeon E7420 quadcore processor at
2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
MHz, was guest host for the VMs. While the other 4 nodes
were auxiliary servers with a Tesla C1090 GPU each one. 
In the software stack the testbed system operates a
Centos 6.6; the cloud platform is deployed by OpenStack; and
the GPUs use CUDA 6.5. The GPU virtualization support is
based on rCUDA v5.0b.
We have designed 6 different set-ups which can be divided
in 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
“Exclusive” mode will provide, at most, 4 accelerators. Whilst
the number of available GPU in “Shared” mode will depend
on the partition size. In this particular case, we have halved
the memory, so we can use up to 8 GPUs. For each group, we
have deployed virtual clusters of 1, 2 and 3 nodes where will
run the processes of the application. The instances are based
on the flavor “m1.medium” which determines VM of 2 cores
and 4096 MB of memory.
MCUDA-MEME~\cite{Liu2010} has been the application executed with
different configurations in order to test the set-ups. MCUDA-
MEME is GPGPU MPI application, so the number of GPUs
will determine the number of executable processes.
The Figure~\ref{fig3} shows and compares the execution time of
the application with different configurations over different set-
ups. We can determine that with 1 node, the performance is
lower, with 2 and 3 is quite similar, though. Hence, the greatest
difference can be appreciated between modes, where the
highest performance is achieved by sharing the accelerators.

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.5\textwidth]{images/chart.jpg}
  \caption{Execution time of MCUDA-MEME}
  \label{fig3}
\end{figure}

\section{Discussion}
\label{sec:discussion}
-Volver a hacer un resumen
-posibles futuras aplicaciones (lab)
-pruebas con otros proveedores o con otras plataformas (azure)
